---
title: "Prescribing Costs in Hospitals and the Community"
pagetitle: Prescribing Costs in Hospitals and the Community
always_allow_html: TRUE
output:
  html_document:
    anchor_sections: no
    css: www/style.css
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
  pdf_document:
    toc: yes
    toc_depth: '2'
  word_document:
    toc: yes
    toc_depth: '2'
---


<script>
  <!-- $(document).ready(function() { -->
  <!--   $('#TOC').prepend('<div><img src=`r knitr::image_uri("www/national-statistics-logo.svg")` alt=\"National Statistics logo\" style=\"position: relative; width: 50%; margin: 10% auto 10% auto; display: block;\"></div>'); -->
  <!-- }); -->
  
  
   $(document).ready(function() {
    $('#header').prepend('<img src=`r knitr::image_uri("www/bsa_logo.svg")` alt=\"NHSBSA logo\" style=\"position:relative; float: right; top: 0; right: 0; padding: 10px; width: 25%;\">');
  });
  
</script>


<style type="text/css">

* {
font-family: Arial;
}

body, td {
   font-size: 16px;
   font-family: sans-serif;
}

h2 {
color: #005EB8;
}
</style>
<html lang="en">
```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
# set code chunk options to disable echo by default
knitr::opts_chunk$set(echo = FALSE)

```
<main>

# England April 2017 to March 2022 {.toc-ignore}

Published 10 November 2022

## Key findings

When using actual costs in primary care the total expenditure on medicines in England by the NHS in 2021/22 was estimated to be £17.3 billion. The total expenditure without central rebates was £17.8 billion, an increase of 8.32% from the £16.4 billion in 2020/21. This is split between £8.90 billion issued in primary care and dispensed in the community, £8.77 billion issued in hospitals, £85.1 million prescribed in hospitals and dispensed in the community, and £29.3 million prescribed by dentists and dispensed in the community. 

* The cost of medicines prescribed in primary care in England in 2021/22 was £8.90 billion, 50.0% of total expenditure when using actual costs. 

* The cost of medicines dispensed in hospitals in England in 2021/22 was £8.77 billion, 49.3% of total expenditure.

* In 2021/22, the BNF Chapter 8 – Malignant Disease and Immunosuppression had the highest spend across all settings of £3.11 billion when using Net Ingredient Cost (NIC) in primary care, 16.9% of the total spend. Of this spend, £2.95 billion was for medicines issued and dispensed in hospitals.

* The average costs per capita when using NIC in primary care for drugs, dressings, appliances, and medical devices prescribed in England in 2021/22 across all settings was £324. This is an increase from £301 in 2020/21.

It is not possible to include the central rebates in the figures broken down by Integrated Care Board (ICB) or British National Formulary (BNF) category owing to commercial confidentiality. The central rebate figure includes rebates from arrangements NHS England and NHS Improvement (NHSEI) agreed with pharmaceutical companies following successful negotiations to commission a variety of treatments both in the cancer drugs fund (CDF) and in routine commissioning. This includes treatments for both rare diseases and more common conditions.

---

## 1. Introduction

### 1.1 Scope

These statistics cover NHS prescriptions and medicines issued in England. This includes:

*	prescriptions issued by GP practices and community prescribers in England that have been dispensed in the community in the UK (excluding Northern Ireland)

*	prescriptions issued by hospitals in England that have been dispensed in the community in the UK (excluding Northern Ireland)

*	prescriptions issued by dental practitioners that have been dispensed in the community in the UK (excluding Northern Ireland)

*	medicines issued in hospitals in England that have been dispensed via the hospital pharmacy, homecare companies and outsourced out-patient pharmacy partnerships

But excludes:

* Prescriptions issued through Justice and Armed Services health services in England commissioned by NHSEI but not dispensed in the community, this covers pharmacy, appliances, Dispensing Doctors (DD) and Personally Administered Items (PADM).

* Any medicines issued in hospital in England but not managed via the hospital pharmacy service.

This publication is produced using data that reflects the actual costs paid by hospitals, including any discounts. This data is collated by Rx-Info and is accessed via their Define system. Data from this source is only available from April 2017 onwards. Primary care data is extracted from the NHSBSA Enterprise Data Warehouse (EDW).

This publication aims to explore the costs of medicines prescribed within both primary and secondary care in England. These statistics detail:

* cost breakdowns by financial year and month within each setting

* cost breakdowns by BNF Chapter and BNF Section level by financial year and setting

* financial year cost breakdowns in each setting by ICB

* a view of the data down to financial year, BNF Chapter, BNF Section and ICB combined

### 1.2 Definitions

Throughout this publication when 'costs' are referred to within hospitals they are the actual costs paid (including applicable VAT) for drugs, dressing, appliances, and medical devices which have been issued and used in NHS hospitals in England, following any discount. 

<<ACTUAL COSTS DEFINITION HERE>>

Net Ingredient Cost (NIC) in primary care is the basic price given either in the Drug Tariff or is determined from prices published by manufacturers, wholesalers, or suppliers. Basic price is set out in Parts VIII and IX of the Drug Tariff. For any drugs not in Part VIII, the price is usually taken from the manufacturer, wholesaler or supplier of the product. 'Costs' also includes retained margin, where applicable. This retained medicine margin contributes to the provision of pharmaceutical services alongside the other agreed fees and allowances.

These statistics use the therapeutic classifications defined in the BNF to group medicines together based on their primary therapeutic indication. The NHSBSA uses and maintains the classification system of the BNF implemented prior to the release of edition 70, including the six pseudo BNF chapters (18 to 23) created by NHS Prescription Services used to classify products that fall outside of chapters 1 to 15.

ICBs are a statutory NHS organisation responsible for developing a plan in collaboration with NHS trusts/foundation trusts and other system partners for meeting the health needs of the population, managing the NHS budget and arranging for the provision of health services in the defined area. They took over the functions of Sustainability and Transformation Partnerships (STPs) in July 2022.

Prescribing volumes are not included in this publication. Unlike in the primary care dataset, where item level data is available, in secondary care, data is collated as quantities issued. The concept of a prescription item does not exist for secondary care medicines data.

---

## 2. Results and commentary

The charts in this report are interactive and allow you to isolate the topics that you are interested in by selecting or deselecting a series from the legend of the chart. For example, if you only want to view a single series on a chart, you can click the series names  you are not interested in within the legend of the chart. This will remove the item from the charts. If you wish to subsequently include them, just click their legend entry again.

### 2.1 Costs by setting

#### Figure 1: Total costs by setting between 2017/2018 and 2021/2022 using actual costs in primary care
```{r fy_actual_costs, out.width = "100%"}

chart_data <- table_1 |>
  pivot_longer(
    cols = contains("GBP"),
    names_to = "Area",
    values_to = "Cost"
  ) |>
  mutate(
    sf = 
      prettyNum(
        signif(
      Cost, 
      3
    ),
    big.mark = ","
  )
  )

figure_1 <- highchart() |>
  hc_add_series(
    name = "Primary care in the community",
    type = "column",
    data = chart_data |> filter(Area == "Primary care prescribing dispensed in the community (GBP)"),
    color = "#009639",
    hcaes(
      x = `Financial Year`,
      y = Cost,
      sf = sf
    )
  ) |>
  hc_add_series(
    name = "Issued in hospitals",
    type = "column",
    data = chart_data |> filter(Area == "Hospital prescribing issued within hospitals (GBP)"),
    color = "#005EB8",
    hcaes(
      x = `Financial Year`,
      y = Cost,
      sf = sf
    )
  ) |>
  hc_add_series(
    name = "Hospital in the community",
    type = "column",
    data = chart_data |> filter(Area == "Hospital prescribing dispensed in the community (GBP)"),
    color = "#7C2855",
    hcaes(
      x = `Financial Year`,
      y = Cost,
      sf = sf
    )
  ) |>
  hc_add_series(
    name = "Dental in the community",
    type = "column",
    data = chart_data |> filter(Area == "Dental prescribing dispensed in the community (GBP)"),
    color = "#768692",
    hcaes(
      x = `Financial Year`,
      y = Cost,
      sf = sf
    )
  ) |>
  hc_xAxis(
    type = "category",
    title =list(
      text = "Financial year"
    )
  ) |>
  hc_yAxis(
    title =list(
      text = "Costs (GBP)"
    )
  ) |>
  hc_plotOptions(
    column = list(
      stacking = "normal"
    )
  ) |>
  highcharter::hc_credits(enabled = TRUE) |>
  hc_tooltip(
    shared = T,
    pointFormat = '<span style="color:{point.color}">\u25CF</span> {series.name}: <b>\u00A3{point.sf}</b><br/>'  )

figure_1

```

Source: [Summary Tables - Prescribing Costs in Hospitals and the Community - Table One](https://nhsbsa-opendata.s3.eu-west-2.amazonaws.com/pchc/pchc-2022-summary-tables-v001.xlsx)

The actual cost of items prescribed in primary care in England in 2021/22 was £8.90 billion, 50.0% of total expenditure. In 2017/18 the actual cost of items prescribed in primary care in England was £8.26 billion, 54.9% of total expenditure. 

The cost of items dispensed in hospitals in England in 2021/22 was £8.77 billion, 49.3% of total expenditure. In 2017/18 the cost of items dispensed in hospitals in England was £6.63 billion, 44.1% of total expenditure. 

The central rebate accounted for £495 million in financial year 2021/22. This would reduce the overall expenditure in hospitals for 2021/22 to £8.27 billion. This figure has not been accounted for in the charts and summary tables as it is not available at a more granular level than overall amount.

#### Figure 2: Total costs by setting between 2017/2018 and 2021/2022 using Net Ingredient Costs (NIC) in primary care
```{r fy_nic, out.width = "100%"}
chart_data <- table_2 |>
  pivot_longer(
    cols = contains("GBP"),
    names_to = "Area",
    values_to = "Cost"
  ) |>
  mutate(
    sf = 
      prettyNum(
        signif(
      Cost, 
      3
    ),
    big.mark = ","
  )
  )

figure_2 <- highchart() |>
  hc_add_series(
    name = "Primary care in the community",
    type = "column",
    data = chart_data |> filter(Area == "Primary care prescribing dispensed in the community (GBP)"),
    color = "#009639",
    hcaes(
      x = `Financial Year`,
      y = Cost,
      sf = sf
    )
  ) |>
  hc_add_series(
    name = "Issued in hospitals",
    type = "column",
    data = chart_data |> filter(Area == "Hospital prescribing issued within hospitals (GBP)"),
    color = "#005EB8",
    hcaes(
      x = `Financial Year`,
      y = Cost,
      sf = sf
    )
  ) |>
  hc_add_series(
    name = "Hospital in the community",
    type = "column",
    data = chart_data |> filter(Area == "Hospital prescribing dispensed in the community (GBP)"),
    color = "#7C2855",
    hcaes(
      x = `Financial Year`,
      y = Cost,
      sf = sf
    )
  ) |>
  hc_add_series(
    name = "Dental in the community",
    type = "column",
    data = chart_data |> filter(Area == "Dental prescribing dispensed in the community (GBP)"),
    color = "#768692",
    hcaes(
      x = `Financial Year`,
      y = Cost,
      sf = sf
    )
  ) |>
  hc_xAxis(
    type = "category",
    title =list(
      text = "Financial year"
    )
  ) |>
  hc_yAxis(
    title =list(
      text = "Costs (GBP)"
    )
  ) |>
  hc_plotOptions(
    column = list(
      stacking = "normal"
    )
  ) |>
  highcharter::hc_credits(enabled = TRUE) |>
  hc_tooltip(
    shared = T,
    pointFormat = '<span style="color:{point.color}">\u25CF</span> {series.name}: <b>\u00A3{point.sf}</b><br/>'  )

figure_2

```

Source: [Summary Tables - Prescribing Costs in Hospitals and the Community - Table Two](https://nhsbsa-opendata.s3.eu-west-2.amazonaws.com/pchc/pchc-2022-summary-tables-v001.xlsx)

The NIC of items prescribed in primary care in England in 2021/22 was £9.49 billion, 51.6% of total expenditure. In 2017/18 the actual cost of items prescribed in primary care in England was £8.88 billion, 56.6% of total expenditure. 


#### Figure 3: Percentage change in costs from previous financial year by setting between between 2017/2018 and 2021/2022 using Net Ingredient Costs (NIC) in primary care
```{r fy_changes, out.width = "100%"}

chart_data <- table_3 |>
  pivot_longer(
    cols = contains("%"),
    names_to = "Area",
    values_to = "Cost"
  ) |>
  mutate(
    sf = 
      prettyNum(
        round(signif(
      Cost, 
      3
    ),2),
    big.mark = ","
  )
  )

figure_3 <- highchart() |>
  hc_add_series(
    name = "Primary care in the community",
    type = "line",
    data = chart_data |> filter(Area == "Primary care prescribing dispensed in the community (%)"),
    color = "#009639",
    hcaes(
      x = `Financial Year`,
      y = Cost,
      sf = sf
    )
  ) |>
  hc_add_series(
    name = "Issued in hospitals",
    type = "line",
    data = chart_data |> filter(Area == "Hospital prescribing issued within hospitals (%)"),
    color = "#005EB8",
    hcaes(
      x = `Financial Year`,
      y = Cost,
      sf = sf
    )
  ) |>
  hc_add_series(
    name = "Hospital in the community",
    type = "line",
    data = chart_data |> filter(Area == "Hospital prescribing dispensed in the community (%)"),
    color = "#7C2855",
    hcaes(
      x = `Financial Year`,
      y = Cost,
      sf = sf
    )
  ) |>
  hc_add_series(
    name = "Dental in the community",
    type = "line",
    data = chart_data |> filter(Area == "Dental prescribing dispensed in the community (%)"),
    color = "#768692",
    hcaes(
      x = `Financial Year`,
      y = Cost,
      sf = sf
    )
  ) |>
  hc_xAxis(
    type = "category",
    title =list(
      text = "Financial year"
    )
  ) |>
  hc_yAxis(
    title =list(
      text = "Change in costs (%)"
    )
  ) |>
  hc_plotOptions(
    column = list(
      stacking = "normal"
    )
  ) |>
  hc_tooltip(
    shared = T,
    pointFormat = '<span style="color:{point.color}">\u25CF</span> {series.name}: <b>{point.sf}%</b><br/>'
  ) |>
  highcharter::hc_credits(enabled = TRUE)

figure_3

```

Source: [Summary Tables - Prescribing Costs in Hospitals and the Community - Table Three](https://nhsbsa-opendata.s3.eu-west-2.amazonaws.com/pchc/pchc-2022-summary-tables-v001.xlsx)

The cost of items within primary care in the community in England in 2021/22 was £9.49 billion. This was a 0.74% increase of £70.0 million from £9.42 billion in 2020/21 and a 6.93% increase of £616 million from £8.88 billion in 2017/18.

The cost of items dispensed in hospitals in England in 2021/22 was £8.77 billion. This was a 16.8% increase of £1.26 billion from £7.51 billion in 2020/21 and a 32.2% increase of £2.13 billion from £6.63 billion in 2017/18.

In 2021/22 the total costs paid for drugs, dressing, appliances, and medical devices was £18.4 billion. This was an increase of 7.86% from the £17.0 billion paid in 2020/21 and an increase of 15.9% from the £15.7 billion paid in 2017/18.

#### Figure 4: Total costs by setting between April 2017 and March 2022 using Net Ingredient Costs (NIC) in primary care
```{r monthly_summary, out.width = "100%"}

chart_data <- table_4 |>
  pivot_longer(
    cols = contains("GBP"),
    names_to = "Area",
    values_to = "Cost"
  ) |>
  mutate(
    sf = 
      prettyNum(
        round(signif(
      Cost, 
      3
    ),2),
    big.mark = ","
  ),
  `Year Month` = as.Date(
    paste(`Year Month`,"01",sep=""),
    format = "%Y%m%d"
  )
  )

figure_4 <- highchart() |>
  hc_add_series(
    name = "Primary care in the community",
    type = "line",
    data = chart_data |> filter(Area == "Primary care prescribing dispensed in the community (GBP)"),
    color = "#009639",
    hcaes(
      x = `Year Month`,
      y = Cost,
      sf = sf
    )
  ) |>
  hc_add_series(
    name = "Issued in hospitals",
    type = "line",
    data = chart_data |> filter(Area == "Hospital prescribing issued within hospitals (GBP)"),
    color = "#005EB8",
    hcaes(
      x = `Year Month`,
      y = Cost,
      sf = sf
    )
  ) |>
  hc_add_series(
    name = "Hospital in the community",
    type = "line",
    data = chart_data |> filter(Area == "Hospital prescribing dispensed in the community (GBP)"),
    color = "#7C2855",
    hcaes(
      x = `Year Month`,
      y = Cost,
      sf = sf
    )
  ) |>
  hc_add_series(
    name = "Dental in the community",
    type = "line",
    data = chart_data |> filter(Area == "Dental prescribing dispensed in the community (GBP)"),
    color = "#768692",
    hcaes(
      x = `Year Month`,
      y = Cost,
      sf = sf
    )
  ) |>
  hc_xAxis(
    type = "datetime",
    title =list(
      text = "Month"
    )
  ) |>
  hc_yAxis(
    title =list(
      text = "Costs (GBP)"
    )
  ) |>
  hc_plotOptions(
    line = list(
      marker = list(
        enabled = FALSE
      )
    )
  ) |>
  hc_tooltip(
    shared = T,
    pointFormat = '<span style="color:{point.color}">\u25CF</span> {series.name}: <b>\u00A3{point.sf}</b><br/>'
  ) |>
  highcharter::hc_credits(enabled = TRUE)

figure_4

```

Source: [Summary Tables - Prescribing Costs in Hospitals and the Community - Table Four](https://nhsbsa-opendata.s3.eu-west-2.amazonaws.com/pchc/pchc-2021-summary-tables-v001.xlsx)

In March 2022, the total actual costs paid for drugs, dressing, appliances, and medical devices which have been issued and used in NHS hospitals in England stood at £814 million. This in an increase of 11.8% on the £728 million paid in March 2021 and an increase of 40.8% on the £578 million paid in March 2018. 

### 2.2 Costs by BNF Chapter

#### Figure 5: Total costs by BNF Chapter in the 2021/2022 financial year using Net Ingredient Costs (NIC) in primary care
```{r bnf_summary, out.width = "100%"}


chart_data <- table_5 |>
  pivot_longer(
    cols = contains("GBP"),
    names_to = "Area",
    values_to = "Cost"
  ) |>
  mutate(
    sf = 
      prettyNum(
        round(signif(
      Cost, 
      3
    ),2),
    big.mark = ","
  )
  ) |>
  filter(
    `Financial Year` == max(`Financial Year`)
  )

figure_5 <- highchart() |>
  hc_add_series(
    name = "Primary care in the community",
    type = "column",
    data = chart_data |> filter(Area == "Primary care prescribing dispensed in the community (GBP)"),
    color = "#009639",
    hcaes(
      x = `BNF Chapter Code`,
      y = Cost,
      sf = sf,
      desc = `BNF Chapter Name`
    )
  ) |>
  hc_add_series(
    name = "Issued in hospitals",
    type = "column",
    data = chart_data |> filter(Area == "Hospital prescribing issued within hospitals (GBP)"),
    color = "#005EB8",
    hcaes(
      x = `BNF Chapter Code`,
      y = Cost,
      sf = sf,
      desc = `BNF Chapter Name`
    )
  ) |>
  hc_add_series(
    name = "Hospital in the community",
    type = "column",
    data = chart_data |> filter(Area == "Hospital prescribing dispensed in the community (GBP)"),
    color = "#7C2855",
    hcaes(
      x = `BNF Chapter Code`,
      y = Cost,
      sf = sf,
      desc = `BNF Chapter Name`
    )
  ) |>
  hc_add_series(
    name = "Dental in the community",
    type = "column",
    data = chart_data |> filter(Area == "Dental prescribing dispensed in the community (GBP)"),
    color = "#768692",
    hcaes(
      x = `BNF Chapter Code`,
      y = Cost,
      sf = sf,
      desc = `BNF Chapter Name`
    )
  ) |>
  hc_yAxis(
    title =list(
      text = "Costs (GBP)"
    )
  ) |>
  hc_xAxis(
    type = "category",
    title =list(
      text = "BNF Chapter"
    )
  ) |>
  hc_plotOptions(
    column = list(
      stacking = "normal"
    )
  ) |>
  hc_tooltip(
    shared = T,
    headerFormat = '{point.key}: {point.point.desc}<br>',
    pointFormat = '<span style="color:{point.color}">\u25CF</span> {series.name}: <b>\u00A3{point.sf}</b><br/>'
  ) |>
  highcharter::hc_credits(enabled = TRUE)

figure_5

```

Source: [Summary Tables - Prescribing Costs in Hospitals and the Community - Table Five](https://nhsbsa-opendata.s3.eu-west-2.amazonaws.com/pchc/pchc-2022-summary-tables-v001.xlsx)

In 2021/22 the largest single spend for a BNF Chapter in any setting was for medicines in chapter 8 - Malignant Disease and Immunosuppression issued in hospitals at £2.95 billion. This represents 33.6% of the total spend within hospitals. This is also the BNF Chapter with the highest costs across all combined settings in 2020/21 at £3.11 billion, 16.9% of the total spend of £18.4 billion.

Within primary care in the community in 2021/22 chapter 2 - Cardiovascular System accounted for the largest spend at £1.61 billion. This was 17.0% of the total costs in primary care prescribing dispensed in the community.

Within hospitals in 2021/22 there was a spend of £382 million which is not accounted for under any of the BNF Chapters. These are drugs manufactured as specials, dressings, appliances, and medical devices which are not authored on the Dictionary of Medicines and Devices (dm+d).

#### Figure 6: Top 10 BNF Sections by spend across all settings in the 2021/2022 financial year using Net Ingredient Costs (NIC) in primary care
```{r bnf_section_summary, out.width = "100%"}

chart_data_top_n <- table_6 |>
  pivot_longer(
    cols = contains("GBP"),
    names_to = "Area",
    values_to = "Cost"
  ) |>
  mutate(
    sf = 
      prettyNum(
        round(signif(
      Cost, 
      3
    ),2),
    big.mark = ","
  )
  ) |>
  filter(
    `Financial Year` == max(`Financial Year`)
  ) |>
  filter(
    Area == "Total (GBP)"
  ) |>
  top_n(
    10,
    Cost
  ) |>
  arrange(
    desc(Cost)
  )


chart_data <- table_6 |>
  pivot_longer(
    cols = contains("GBP"),
    names_to = "Area",
    values_to = "Cost"
  ) |>
  mutate(
    sf = 
      prettyNum(
        round(signif(
          Cost, 
          3
        ),2),
        big.mark = ","
      )
  ) |>
  filter(
    `Financial Year` == max(`Financial Year`),
    `BNF Section Code` %in% chart_data_top_n$`BNF Section Code`
  ) |>
  arrange(match(`BNF Section Code`, c(chart_data_top_n$`BNF Section Code`)))

figure_6 <- highchart() |>
  hc_add_series(
    name = "Primary care in the community",
    type = "column",
    data = chart_data |> filter(Area == "Primary care prescribing dispensed in the community (GBP)"),
    color = "#009639",
    hcaes(
      x = `BNF Section Code`,
      y = Cost,
      sf = sf,
      desc = `BNF Section Name`
    )
  ) |>
  hc_add_series(
    name = "Issued in hospitals",
    type = "column",
    data = chart_data |> filter(Area == "Hospital prescribing issued within hospitals (GBP)"),
    color = "#005EB8",
    hcaes(
      x = `BNF Section Code`,
      y = Cost,
      sf = sf,
      desc = `BNF Section Name`
    )
  ) |>
  hc_add_series(
    name = "Hospital in the community",
    type = "column",
    data = chart_data |> filter(Area == "Hospital prescribing dispensed in the community (GBP)"),
    color = "#7C2855",
    hcaes(
      x = `BNF Section Code`,
      y = Cost,
      sf = sf,
      desc = `BNF Section Name`
    )
  ) |>
  hc_add_series(
    name = "Dental in the community",
    type = "column",
    data = chart_data |> filter(Area == "Dental prescribing dispensed in the community (GBP)"),
    color = "#768692",
    hcaes(
      x = `BNF Section Code`,
      y = Cost,
      sf = sf,
      desc = `BNF Section Name`
    )
  ) |>
   hc_xAxis(
     type = "category",
     title =list(
      text = "BNF Section"
     )
   ) |>
  hc_yAxis(
    title =list(
      text = "Costs (GBP)"
    )
  ) |>
  hc_plotOptions(
    column = list(
      stacking = "normal"
    )
  ) |>
  hc_tooltip(
    shared = T,
    headerFormat = '{point.key}: {point.point.desc}<br>',
    pointFormat = '<span style="color:{point.color}">\u25CF</span> {series.name}: <b>\u00A3{point.sf}</b><br/>'
  ) |>
  highcharter::hc_credits(enabled = TRUE)

figure_6

```

Source: [Summary Tables - Prescribing Costs in Hospitals and the Community - Table Six](https://nhsbsa-opendata.s3.eu-west-2.amazonaws.com/pchc/pchc-2022-summary-tables-v001.xlsx)

In 2021/22 the largest single spend for a BNF Section in any setting was for medicines in section 0801 - Cytotoxic Drugs issued in hospitals at £1.98 billion. This represents 22.6% of the total spend within hospitals.

Within primary care in the community in 2021/22 section 0601 - Drugs Used in Diabetes accounted for the largest spend at £1.25 billion. This was 13.2% of the total costs in primary care prescribing dispensed in the community.

### 2.3 Costs by ICB

#### Figure 7: Total costs by ICB in the 2021/2022 financial year using Net Ingredient Costs (NIC) in primary care
```{r stp_summary, out.width = "100%"}

chart_data_order <- table_7 |>
  pivot_longer(
    cols = contains("GBP"),
    names_to = "Area",
    values_to = "Cost"
  ) |>
  filter(
    `Financial Year` == max(`Financial Year`)
  ) |>
  filter(
    Area == "Total (GBP)"
  ) |>
  arrange(
    desc(Cost)
  )


chart_data <- table_7 |>
  pivot_longer(
    cols = contains("GBP"),
    names_to = "Area",
    values_to = "Cost"
  ) |>
  mutate(
    sf = 
      prettyNum(
        round(signif(
          Cost, 
          3
        ),2),
        big.mark = ","
      )
  ) |>
  filter(
    `Financial Year` == max(`Financial Year`)
  ) |>
  arrange(match(ICB, c(chart_data_order$ICB)))

figure_7 <- highchart() |>
  hc_add_series(
    name = "Primary care in the community",
    type = "column",
    data = chart_data |> filter(Area == "Primary care prescribing dispensed in the community (GBP)"),
    color = "#009639",
    hcaes(
      x = `ICB Code`,
      y = Cost,
      sf = sf,
      icb = `ICB`
    )
  ) |>
  hc_add_series(
    name = "Issued in hospitals",
    type = "column",
    data = chart_data |> filter(Area == "Hospital prescribing issued within hospitals (GBP)"),
    color = "#005EB8",
    hcaes(
      x = `ICB Code`,
      y = Cost,
      sf = sf,
      icb = `ICB`
    )
  ) |>
  hc_add_series(
    name = "Hospital in the community",
    type = "column",
    data = chart_data |> filter(Area == "Hospital prescribing dispensed in the community (GBP)"),
    color = "#7C2855",
    hcaes(
      x = `ICB Code`,
      y = Cost,
      sf = sf,
      icb = `ICB`
    )
  ) |>
  hc_xAxis(
    type = "category",
    title =list(
      text = "ICB"
    )
  ) |>
  hc_yAxis(
    title =list(
      text = "Costs (GBP)"
    )
  ) |>
  hc_plotOptions(
    column = list(
      stacking = "normal"
    )
  ) |>
  hc_tooltip(
    shared = T,
    headerFormat = '{point.key}: {point.point.icb}<br>',
    pointFormat = '<span style="color:{point.color}">\u25CF</span> {series.name}: <b>\u00A3{point.sf}</b><br/>'
  ) |>
  highcharter::hc_credits(enabled = TRUE)

figure_7

```

Source: [Summary Tables - Prescribing Costs in Hospitals and the Community - Table Seven](https://nhsbsa-opendata.s3.eu-west-2.amazonaws.com/pchc/pchc-2022-summary-tables-v001.xlsx)

NHS Greater Manchester was the ICB with the highest costs across all settings of £1.11 billion in 2021/22. 

Data under the 'Undefined' ICB relate to costs in both the primary and secondary care datasets which we have been unable to map to an ICB.

Data around dental in the community has not been included in figure 5 as dental data is captured under a national code and not available at ICB level.

ICB costs in secondary care are based on the ICB in which the hospital is located. Not all spend is specifically for that ICB (e.g. patients may be referred to a hospital outside of their usual ICB of residence).

#### Figure 8: Average costs per capita by ICB in the 2021/2022 financial year using Net Ingredient Costs (NIC) in primary care
```{r stp_per_capita, out.width = "100%"}

chart_data <- table_7 |>
  pivot_longer(
    cols = contains("GBP"),
    names_to = "Area",
    values_to = "Cost"
  ) |>
  mutate(
    sf = 
      prettyNum(
        round(signif(
          Cost, 
          3
        ),2),
        big.mark = ","
      )
  ) |>
  filter(
    `Financial Year` == max(`Financial Year`),
    Area == "Average Cost Per Capita (GBP)"
  ) |>
  filter(`ICB Code` != "-") |>
  arrange(desc(Cost)) |>
  mutate(total_cost = Cost * Population)

ave_per_capita <- sum(chart_data$total_cost, na.rm = T)/sum(chart_data$Population, na.rm = T)

chart_data <- chart_data |>
  mutate(ave = ave_per_capita,
         sf_ave = prettyNum(as.numeric(signif(ave,3)), big.mark = ","))

figure_8 <- highchart() |>
  hc_add_series(
    name = "ICB",
    type = "column",
    data = chart_data,
    color = "#005EB8",
    hcaes(
      x = `ICB Code`,
      y = `Cost`,
      sf = sf,
      icb = ICB
    )
  ) |>
  hc_add_series(
    name = "National average",
    type = "line",
    data = chart_data,
    color = "black",
    hcaes(
      x = `ICB Code`,
      y = ave,
      sf = sf_ave
    )
  ) |>
  hc_xAxis(
    type = "category",
    title = list(
      text = "ICB"
    )
  ) |>
  hc_yAxis(
    title =list(
      text = "Costs (GBP)"
    )
  ) |>
  hc_plotOptions(
    column = list(
      stacking = "normal"
    ),
    line = list(
      marker = list(
        enabled = FALSE
      )
    )
  ) |>
  hc_tooltip(
    shared = T,
    headerFormat = '{point.key}: {point.point.icb}<br>',
    pointFormat = '<span style="color:{point.color}">\u25CF</span> {series.name}: <b>\u00A3{point.sf}</b><br/>'
  ) |>
  highcharter::hc_credits(enabled = TRUE)

figure_8

```

Source: [Summary Tables - Prescribing Costs in Hospitals and the Community - Table Seven](https://nhsbsa-opendata.s3.eu-west-2.amazonaws.com/pchc/pchc-2022-summary-tables-v001.xlsx)

The average cost per capita for medicines issued in all settings in England in 2021/22 was £324. This is an increase from £301 in 2020/21 and an increase from £274 in 2017/18.

NHS Birmingham and Solihull was the ICB with the largest average per capita costs in 2021/22 at £516. NHS Hertfordshire and West Essex was the lowest at £224. 

---

## 3. Changes to this publication

We have added an additional table and figure which now includes the actual costs in primary care. A definition of actual costs can be found in section 1.2 of this publication.

In July 2022 Integrated Care Boards (ICBs) replaced Sustainability and Transformation Plans (STPs) as the legal organisations responsible for commissioning pharmaceutical services. In these statistics we have used the latest NHS organisational structure.

For full details of the changes made please refer to the [Background Information and Methodology](https://nhsbsa.nhs.uk/sites/default/files/2021-11/PCHC_2021_background_information_methodology_v001.pdf) note released alongside these statistics.

---

## 4. About these statistics

### 4.1 Uses of these statistics

These statistics have a variety of uses including:

* Obtaining a national view of costs for prescriptions dispensed in hospitals and the community in England across the financial year.

* Analysis of cost trends by BNF Chapter, BNF Section and IBC.

* Allowing public scrutiny of national prescribing habits.

### 4.2 British National Formulary (BNF) 

The primary care data used in this publication comes from the NHSBSA Enterprise Data Warehouse (EDW). This includes the split of the data down to BNF Section. The data for secondary care is provided by Rx-Info. Rx-Info use an extract from the dm+d to obtain BNF classifications.

### 4.3 Time periods reported in this publication

From the 2021/22 publication onward the starting point has switched to April 2017. Secondary care data provided by Rx-Info is only available back to this point. As the RX-Info Define database only holds data from up to 5 years ago, this publication will present a rolling 5 year timescale.

### 4.4 Geographies included in this publication

Integrated Care Boards (ICB) are health geographies that are formed by NHS organisations, local councils, and others to set out proposals to improve health and care for the local population. 

In primary care data the ICB geographies shown use NHSBSA administrative records, not geographical boundaries, and more closely reflect the operational organisation of prescribing organisations than other geographical data sources. Within the secondary care data ICB is provided by Rx-Info who use the NHS Digital Organisation Data Service (ODS) as their source.

### 4.5 Population statistics included in this publication

Populations by ICB throughout this report are the same for each year and come from the latest [ONS population estimates](https://www.ons.gov.uk/file?uri=%2fpeoplepopulationandcommunity%2fpopulationandmigration%2fpopulationestimates%2fdatasets%2fclinicalcommissioninggroupmidyearpopulationestimates%2fmid2020sape23dt6a/sape23dt6amid2020ccg2021estimatesunformatted.xlsx).

### 4.6 Measures of average

Per capita costs measure the average costs across all settings per person in a given area (England, ICB) in the specified year. It is calculated by dividing the area's total costs by its total population.

### 4.7 Planned changes to this publication

This is an experimental official statistic release. Experimental statistics are newly developed or innovative statistics. These are published so that users and stakeholders can be involved in the assessment of their suitability and quality at an early stage. More information on how users can give us their feedback can be found in section 6 of this summary. We will regularly be reviewing the methodology used within the statistics.

### 4.8 Monthly SCMD data

NHSBSA Information Services also release a monthly [data set](https://opendata.nhsbsa.net/dataset/secondary-care-medicines-data-indicative-price) for Secondary Care Medicines Data via the Open Data Portal (ODP). The ODP is the platform where we host our open data products. The monthly data set is produced from data provided by Rx-Info. This data includes indicative costs rather than actual costs, these are derived from current medicines pricing data held in NHSBSA data systems including the Common Drug Reference (CDR) database and dm+d calculated to a Virtual Medicinal Product (VMP) level.

---

## 5. Accessibility

### How to request content in an accessible format {.toc-ignore}

If you need information on this website in a different format like accessible PDF, large print, easy read, audio recording or braille, you can contact us by:

**Email**: nhsbsa.accessibility@nhs.net

**Phone**: 0191 203 5318

[Find out about call charges](https://www.nhsbsa.nhs.uk/contact-us/call-charges-and-phone-numbers)

We’ll consider your request and get back to you in 5 working days.

These contact details are only for accessibility queries. This inbox is not for technical queries or IT problems. If you have a query that is not about accessibility, go to the ‘Contact us’ section of this page.

View our [Accessibility statement for Official Statistics Narratives](https://www.nhsbsa.nhs.uk/accessibility-statement-official-statistics-narratives).

---

## 6. Feedback and contact us

We invited feedback on the changes made to the Prescribing Costs in Hospitals and the Community Official Statistics publication, including the change of data source. The consultation took place between 15 November 2021 and 4 February 2022. You can find details of this consultation on our [website](https://www.nhsbsa.nhs.uk/announcements/consultations-and-user-engagement/prescription-costs-hospitals-and-community-pchc-consultation).

Feedback is important to us; we welcome any questions and comments relating to these statistics. 

You can complete a [short survey about this publication](https://online1.snapsurveys.com/Official_Statistics_Feedback) to help us improve the Official Statistics that we produce. All responses will remain anonymous and individuals will not be identifiable in any report that we produce.

You can view our [privacy policy](https://www.nhsbsa.nhs.uk/our-policies/privacy) on our website to see how your data is used and stored.

You can contact us by:

**Email:** nhsbsa.statistics@nhs.net

**You can also write to us at:**

NHSBSA - Statistics  
NHS Business Services Authority  
Stella House  
Goldcrest Way  
Newburn Riverside  
Newcastle upon Tyne  
NE15 8NY

**Responsible statistician:** Graham Platten

</main>